Combinatorial Proofs and Decomposition Theorems for First-order Logic\vspace{-3ex}

### 2021-01-25 main.tex
Minor fixes and typos

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
  We uncover a close relationship between combinatorial and
  syntactic proofs for first-order logic. Whereas syntactic proofs are
  formalized in a deductive proof system based on inference rules,
  a combinatorial proof is a syntax-fre presentation of a proof
  that is independent from any set of inference rules. We show that
  the two proof representations are related via a deep inference
  decomposition theorem that establishes a new kind of normal form
  for syntactic proofs. This yields (a) a simple proof
  of soundness and completeness for first-order combinatorial
  proofs, and (b) a full completeness theorem: every combinatorial
  proof is the image of a syntactic proof.
\end{abstract}

@@

For proving this theorem, we have to show that no inference rule in
$\FOMLS$ can introduce a bimatching. To simplify the argument, we
introduce the \bfit{frame}~\cite{hughes:unifn} of the linked fograph $\gC$,
which is a linked (proposional) cograph in which the dependencies
between the binders in $\gC$ are encoded as links.

@@

\begin{proof}[Proof of Theorem~\ref{thm:MLS1->fonet}]
  From $\Deri$ we construct a derivation $\frameof\Deri$ of $\frameof
  A$ in the propositional fragment of $\FOMLS$, such that
  \hbox{$\fographof{\frameof\Deri}=\frameof{\fographof\Deri}$}.  The
  rules $\aiD,\tttD,\mix$ and $\sw$ are translated trivially, and for
  $\fequ$, it suffices to observe that the frame construction is
  invariant under $\fequ$. Finally, for the rules $\forall$ and
  $\exists$, proceed as follows. Every instance of $\forall$ is
  replaced by the derivation on the right below:\footnote{For better
  readability we omit superfluous parentheses, knowing that we always
  have $\fequ$ incorporating associativity and commutativity of
  $\vlan$ and $\vlor$.} \vadjust{\vskip-2ex}
  \begin{equation*}
    \scalebox{.95}{$
    \vlinf{\forall}{}{\Scons{\forall y_j.\ttt}}{\Scons{\ttt}}
    %\qquad
    \;\leadsto
    %\qquad
    \vlderivation{
      \vlde{\set{\sw,\fequ}}{\DDeri_2}{\Scons{q_{h_1,j}\vlor\cdots\vlor
          q_{h_n,j}\vlor(\dual q_{h_1,j}\vlan\cdots\vlan\dual
          q_{h_n,j}\vlan\ttt)}}{
        \vlde{\set{\aiD,\tttD}}{\DDeri_1}{
          \Scons{(q_{h_1,j}\vlor \dual q_{h_1,j})\vlan\cdots\vlan(q_{h_n,j}\vlor \dual q_{h_n,j})\vlan\ttt}}{
          \vlhy{\ttt}}}}
    $}
  \end{equation*}
  where $h_1,\ldots,h_n$ range over the indices of the existential
  binders dependend on that $y_j$. It is easy to see how $\DDeri_1$ is
  constructed, and for $\DDeri_2$ see,
  e.g. \cite{ATS:esslli2019,gug:str:01,brunnler:tiu:01,dissvonlutz}. Then, every
  occurrence of $\forall y_j.F$ is replaced by
  $q_{h_1,j}\vlor\cdots\vlor q_{h_j,j}\vlor (\dual
  q_{h_1,j}\vlan\cdots\vlan\dual q_{h_j,j}\vlan F)$ in the derivation
  below that $\forall$-instance. Now, observe that all instances of
  the $\exists$-rule introducing $x_i$ dependend on $y_j$ must occur
  below in the derivation (otherwise $\Deri$ would not be rectified).
  Now consider such an instance $\vlinf{\exists}{}{\Scons{\exists
      x_i.B}}{\Scons{B\ssubst {x_i} t}}$.  Its context $\Sconhole$
  must contain all the $\forall y_j$ the $\exists x_i$ depends on,
  such that $B$ is in their scope. Following the translation of the
  $\forall$ rules above, we can therefore translate the $\exists$-rule
  instance by the following derivation
  \begin{equation*}
    \vlder{\set{\sw,\fequ}}{\DDeri_3}{ S_0\cons{S_1\cons{\cdots
          S_{k_l-1}\cons{S_{k_l}\cons{\dual q_{i,k_1}\vlan\dual
              q_{i,k_2}\vlan\cdots q_{i,k_l}\vlan B'}}\cdots}}}{
      S_0\cons{\dual q_{i,k_1}\vlan S_1\cons{\dual
          q_{i,k_2}\vlan\cdots S_{k_l-1}\cons{\dual q_{i,k_k}\vlan
            S_{k_l}\cons{B'}}\cdots}}}
  \end{equation*}
  where $k_1,\ldots,k_l$ are the indices of of the universal binders
  on which that $x_i$ depends, and $B'$ is $B$ in which all predicates
  are replaced by a nullary one (step~3 in the frame construction). The
  derivation $\DDeri_3$ can be constructed in the same way as
  $\DDeri_2$.

  Doing this to all instances of the rules $\forall$ and $\exists$ in
  $\Deri$ yields indeed a propositional derivation $\frameof\Deri$
  with
  \hbox{$\fographof{\frameof\Deri}=\frameof{\fographof\Deri}$}. It has
  been shown by Retor\'e~\cite{retore:pomset:RR} and
  rediscovered by Stra{\ss}burger~\cite{dissvonlutz} that
  $\fographof{\frameof\Deri}=\tuple{\graphof{\frameof
      C},\linkingof{\frameof\Deri}}$ can not contain an induced
  bimatching. By Lemma~\ref{lem:frame}, $\fographof{\Deri}$ does not
  have an induced bimatching either. Furthemore, it followed from the
  definition of $\dsubstof\Deri$ that it is a dualizer for
  $\fographof{\Deri}$.
  %Hence $\fographof{\Deri}$ is a fonet.
  %% We have
  %% to show that no inference rule in $\FOMLS$ can introduce a
  %% bimatching. For the rules $\forall$, $\aiD$, $\tttD$, $\fequ$, and
  %% $\mix$, this is immediate, as they do not introduce new edges in the
  %% cograph connecting two vertices that are each incident to a leap
  %% edge. The only rules that can do that are $\sw$ and $\exists$. For
  %% the $\sw$-rule, we refer reader to~\cite{retore:???} or
  %% \cite{gug:SIS} or \cite{dissvonlutz} where this property has been
  %% shown in detail. For the rule $\exists$, a potential bimatching can
  %% only be created by a dependency between the \todo{need frame?}
\end{proof}

@@

\begin{proof}
   By \cite[Proposition~7.6.1]{str:07:RTA}, we have a derivation
   $\vlder{\set{\wrD,\PE\fequ}}{\DDeri}{\PE H}{\PE G}$. In order to
   lift $\DDeri$, we might need to reorganize the instancese of
   $\wD$. If $H$ contains a subformula $\forall x.A$ which is not
   present in $G$, the $\wD$-instances in $\DDeri$ could introduce the
   parts of the propositional encoding $x\vlor A$ independently.  We
   say that an instance $\rr_1$ of $\wD$ in $\Deri$ is \emph{in the
   scope} of an instance $\rr_2$ of $\wD$ if $\rr_1$ introduced
   formulas that contain a free variable $x$ (i.e., $x$ occurs in a
   term in a predicate) and $\rr_2$ introduces the atom $x$ as a
   subformula (i.e. the propositional encoding of the binder $x$).  We
   can now permute the $\wD$-instances in $\DDeri$ such that whenever
   a rule instance $\rr_1$ depends on an instance $\rr_2$, then
   $\rr_2$ occurs below $\rr_1$ in $\DDeri$.
   We can now lift $\DDeri$ stepwise. First, observe that each line of
   in $\DDeri$ is $\PE\fequ$-equivalent to the propositional encoding $\PE P$ of a
   first-order formula $P$.  We now have to show
   that each other instance of $\wrD$ in $\DDeri$ is indeed the image of a correct
   application of $\wD$ or $\wfaD$ in first-order logic.
   If we have a $\wD$
   \begin{equation*}
     \vlinf{\wD}{}{\PE S\cons{x\vlor \PE A}}{\PE S\cons{\PE A}}
     \qquor
     \vlinf{\wD}{}{\PE S\cons{(x\vlor \PE B)\vlor \PE A}}{\PE S\cons{\PE A}}
   \end{equation*}
   then $x$ cannot occur freely in $A$, as otherwise the fibration
   property would be violated. We can therefore lift these instances to
   \begin{equation*}
     \vlinf{\wfaD}{}{S\cons{\forall x. A}}{S\cons{A}}
     \qquor
     \vlinf{\wD}{}{S\cons{(\forall x. B)\vlor A}}{S\cons{A}}
   \end{equation*}
   respectively.     
  %% A subformula of $\PE B$ is called \emph{weak} if it has been
  %% introduced by a weakening. All subformulas of a weak formula are
  %% also weak. Two weak subformulas $B'$ and $B''$ of $\PE B$ form a
  %% \emph{weak pair} if $\PE B\fequ\Scons{B'\vlor B''}$ for some context
  %% $\Sconhole$. We can assume without loss of generality that
  %% whenever weak subformulas $B'$ and $B''$ form a weak pair, they have
  %% been introduced by the same instance of $\wrD$ in
  %% $\PE\Deri_1$.\footnote{If $\PE\Deri_1$ is not of that shape, it can
  %% brought into this form by simple rule permutations, and then
  %% collapsing several weakenings into a single one.}  Now we show that
  %% $\PE\Deri_1$ can be lifted. For this, observe that whenever a
  %% weakening in $\PE\Deri_1$ deletes an atom $x\in\VAR$, it must also
  %% delete all atoms in the scope of the corresponding quantifier,
  %% because $\phi'$ is a fibration on the binding graph. Hence, each line
  If a  weakening happens inside a subformula $x\vlor \PE C$ or $x\vlan
  \PE C$ in $\DDeri$, then there are the following cases:
  \begin{equation*}%\small
    \scalebox{.9}{$
    \vlinf{\wD}{}{\PE S\cons{x\vlor\PE  D\vlor \PE C}}{\PE S\cons{x\vlor\PE  C}}
    \quad
    \vlinf{\wD}{}{\PE S\cons{x\vlan (\PE D\vlor \PE C)}}{\PE S\cons{x\vlan\PE  C}}
    \quad
    \vlinf{\wD}{}{\PE S\cons{(x\vlor\PE  D)\vlan \PE C}}{\PE S\cons{x\vlan \PE C}}
    $}
  \end{equation*}
  The first two cases can be lifted to
  \begin{equation*}
    \vlinf{\wD}{}{S\cons{\forall x.( D\vlor C)}}{S\cons{\forall x.  C}}
    \quand
     \vlinf{\wD}{}{S\cons{\exists x.( D\vlor C)}}{S\cons{\exists x.  C}}
  \end{equation*}
  respectively. But in the third case,  an $\exists$-quantifier would be
  transformed into an $\forall$-quantifier. But as $\phi$ has to
  preserve existentials, this third case cannot occur. All other situations can be lifted trivially, giving us
  $\vlder{\set{\wrD,\wfaD,\fequ}}{\Deri}{H}{G}$ as desired.
\end{proof}

@@

  \PE{(A\substof\phi)}$ and $\PE B_1$ are both proposional

@@

  proposional encodings, but the whole derivation cannot be

@@

  (and therefore a momomorphism) on the binding graphs, there must be

@@

  \item $\acDx$: We know that the premisse of~\eqref{eq:acDeq} is a

@@

We first study the interactions between two rules. Certain cases are unsolved at
this stage, and they are considered later when we study the interactions between
two non-$\fequ$ rule instances seperated by $\fequ$. Only non-trivial cases are
presented here:
\begin{itemize}
\item $\rr_1/\rr_2$, where $\rr_1 \in \set{\wrD, \wfaD}$ and $\rr_2 \in
\set{\acD, \cfaD, \me, \mfaD, \mexD}$:

@@

\item $\rr_1/\rr_2$, where $\rr_1 \in \set{\acD, \cfaD}$ and $\rr_2 \in
\set{\me, \mfaD, \mexD}$: 
  \begin{equation*}
  \vlderivation{
    \vlin{\mfaD}{}{\Scons{\forall x.(A \vlor B)}}{
      \vlin{\cfaD}{}{\Scons{(\forall x.A) \vlor (\forall x.B)}}{
        \vlhy{\Scons{(\forall x.\forall x.A) \vlor (\forall x.B)}}}}}
  \leadsto
  \vlderivation{
    \vlin{\mfaD}{}{\Scons{\forall x.(A \vlor B)}}{
      \vlin{\fequ}{}{\Scons{(\forall x.A) \vlor (\forall x.B)}}{
        \vlin{\mfaD}{}{\Scons{\forall x.(\forall x.A \vlor B)}}{
          \vlhy{\Scons{(\forall x.\forall x.A) \vlor (\forall x.B)}}}}}}
  \end{equation*}

@@

\begin{equation*}
\vlderivation{
  \vlin{\fequ}{}{\forall y.\forall x.A}{
    \vlin{\wfaD}{(x \notin fv(\forall y.A))}{\forall x.\forall y.A}{
      \vlhy{\forall y.A}}}}
\leadsto
\vlderivation{
  \vlin{\wfaD}{(x \notin fv(A))}{\forall y.\forall x.A}{
    \vlhy{\forall y.A}}}
\end{equation*}

\begin{equation*}
\vlderivation{
  \vlin{\fequ}{}{\forall x.\forall y.A}{
    \vlin{\wfaD}{(x \notin fv(A))}{\forall y.\forall x.A}{
      \vlhy{\forall y.A}}}}
\leadsto
\vlderivation{
  \vlin{\wfaD}{(x \notin fv(\forall y.A))}{\forall x.\forall y.A}{
    \vlhy{\forall y.A}}}
\end{equation*}

@@

  \begin{equation*}
  \vlderivation{
    \vlin{\me}{}{\forall x.((A \vlor B) \vlan (C \vlor D))}{
      \vlin{\fequ}{}{\forall x.((A \vlan C) \vlor (B \vlan
D))}{
        \vlin{\wrD}{}{(B \vlan D) \vlor (\forall x.(A \vlan C))}{
          \vlhy{B \vlan D}}}}}
  \leadsto
  \vlderivation{
    \vlin{\fequ}{}{\forall x.((A \vlor B) \vlan (C \vlor D))}{
      \vlin{\wrD}{}{\forall x.((B \vlor A) \vlan (D \vlor C))}{
        \vlin{\wrD}{}{\forall x.((B \vlor A) \vlan D)}{
          \vlin{\faD}{}{\forall x.(B \vlan D)}{
            \vlhy{B \vlan D}}}}}}
  \end{equation*}
  where in the second case, $x$ is free in $B \vlan D$.

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
  We uncover a close relationship between combinatorial and
  syntactic proofs for first-order logic. Whereas syntactic proofs are
  formalized in a deductive proof system based on inference rules,
  a combinatorial proof is a syntax-free presentation of a proof
  that is independent from any set of inference rules. We show that
  the two proof representations are related via a deep inference
  decomposition theorem that establishes a new kind of normal form
  for syntactic proofs. This yields (a) a simple proof
  of soundness and completeness for first-order combinatorial
  proofs, and (b) a full completeness theorem: every combinatorial
  proof is the image of a syntactic proof.
\end{abstract}

@@

For proving this theorem, we have to show that no inference rule in
$\FOMLS$ can introduce a bimatching. To simplify the argument, we
introduce the \bfit{frame}~\cite{hughes:unifn} of the linked fograph $\gC$,
which is a linked (propositional) cograph in which the dependencies
between the binders in $\gC$ are encoded as links.

@@

\begin{proof}[Proof of Theorem~\ref{thm:MLS1->fonet}]
  From $\Deri$ we construct a derivation $\frameof\Deri$ of $\frameof
  A$ in the propositional fragment of $\FOMLS$, such that
  \hbox{$\fographof{\frameof\Deri}=\frameof{\fographof\Deri}$}.  The
  rules $\aiD,\tttD,\mix$ and $\sw$ are translated trivially, and for
  $\fequ$, it suffices to observe that the frame construction is
  invariant under $\fequ$. Finally, for the rules $\forall$ and
  $\exists$, proceed as follows. Every instance of $\forall$ is
  replaced by the derivation on the right below:\footnote{For better
  readability we omit superfluous parentheses, knowing that we always
  have $\fequ$ incorporating associativity and commutativity of
  $\vlan$ and $\vlor$.} \vadjust{\vskip-2ex}
  \begin{equation*}
    \scalebox{.95}{$
    \vlinf{\forall}{}{\Scons{\forall y_j.\ttt}}{\Scons{\ttt}}
    %\qquad
    \;\leadsto
    %\qquad
    \vlderivation{
      \vlde{\set{\sw,\fequ}}{\DDeri_2}{\Scons{q_{h_1,j}\vlor\cdots\vlor
          q_{h_n,j}\vlor(\dual q_{h_1,j}\vlan\cdots\vlan\dual
          q_{h_n,j}\vlan\ttt)}}{
        \vlde{\set{\aiD,\tttD}}{\DDeri_1}{
          \Scons{(q_{h_1,j}\vlor \dual q_{h_1,j})\vlan\cdots\vlan(q_{h_n,j}\vlor \dual q_{h_n,j})\vlan\ttt}}{
          \vlhy{\ttt}}}}
    $}
  \end{equation*}
  where $h_1,\ldots,h_n$ range over the indices of the existential
  binders dependent on that $y_j$. It is easy to see how $\DDeri_1$ is
  constructed, and for $\DDeri_2$ see,
  e.g. \cite{ATS:esslli2019,gug:str:01,brunnler:tiu:01,dissvonlutz}. Then, every
  occurrence of $\forall y_j.F$ is replaced by
  $q_{h_1,j}\vlor\cdots\vlor q_{h_j,j}\vlor (\dual
  q_{h_1,j}\vlan\cdots\vlan\dual q_{h_j,j}\vlan F)$ in the derivation
  below that $\forall$-instance. Now, observe that all instances of
  the $\exists$-rule introducing $x_i$ dependent on $y_j$ must occur
  below in the derivation (otherwise $\Deri$ would not be rectified).
  Now consider such an instance $\vlinf{\exists}{}{\Scons{\exists
      x_i.B}}{\Scons{B\ssubst {x_i} t}}$.  Its context $\Sconhole$
  must contain all the $\forall y_j$ the $\exists x_i$ depends on,
  such that $B$ is in their scope. Following the translation of the
  $\forall$ rules above, we can therefore translate the $\exists$-rule
  instance by the following derivation
  \begin{equation*}
    \vlder{\set{\sw,\fequ}}{\DDeri_3}{ S_0\cons{S_1\cons{\cdots
          S_{k_l-1}\cons{S_{k_l}\cons{\dual q_{i,k_1}\vlan\dual
              q_{i,k_2}\vlan\cdots q_{i,k_l}\vlan B'}}\cdots}}}{
      S_0\cons{\dual q_{i,k_1}\vlan S_1\cons{\dual
          q_{i,k_2}\vlan\cdots S_{k_l-1}\cons{\dual q_{i,k_k}\vlan
            S_{k_l}\cons{B'}}\cdots}}}
  \end{equation*}
  where $k_1,\ldots,k_l$ are the indices of of the universal binders
  on which that $x_i$ depends, and $B'$ is $B$ in which all predicates
  are replaced by a nullary one (step~3 in the frame construction). The
  derivation $\DDeri_3$ can be constructed in the same way as
  $\DDeri_2$.

  Doing this to all instances of the rules $\forall$ and $\exists$ in
  $\Deri$ yields indeed a propositional derivation $\frameof\Deri$
  with
  \hbox{$\fographof{\frameof\Deri}=\frameof{\fographof\Deri}$}. It has
  been shown by Retor\'e~\cite{retore:pomset:RR} and
  rediscovered by Stra{\ss}burger~\cite{dissvonlutz} that
  $\fographof{\frameof\Deri}=\tuple{\graphof{\frameof
      C},\linkingof{\frameof\Deri}}$ cannot contain an induced
  bimatching. By Lemma~\ref{lem:frame}, $\fographof{\Deri}$ does not
  have an induced bimatching either. Furthermore, it followed from the
  definition of $\dsubstof\Deri$ that it is a dualizer for
  $\fographof{\Deri}$.
  %Hence $\fographof{\Deri}$ is a fonet.
  %% We have
  %% to show that no inference rule in $\FOMLS$ can introduce a
  %% bimatching. For the rules $\forall$, $\aiD$, $\tttD$, $\fequ$, and
  %% $\mix$, this is immediate, as they do not introduce new edges in the
  %% cograph connecting two vertices that are each incident to a leap
  %% edge. The only rules that can do that are $\sw$ and $\exists$. For
  %% the $\sw$-rule, we refer reader to~\cite{retore:???} or
  %% \cite{gug:SIS} or \cite{dissvonlutz} where this property has been
  %% shown in detail. For the rule $\exists$, a potential bimatching can
  %% only be created by a dependency between the \todo{need frame?}
\end{proof}

@@

\begin{proof}
   By \cite[Proposition~7.6.1]{str:07:RTA}, we have a derivation
   $\vlder{\set{\wrD,\PE\fequ}}{\DDeri}{\PE H}{\PE G}$. In order to
   lift $\DDeri$, we might need to reorganize the instances of
   $\wD$. If $H$ contains a subformula $\forall x.A$ which is not
   present in $G$, the $\wD$-instances in $\DDeri$ could introduce the
   parts of the propositional encoding $x\vlor A$ independently.  We
   say that an instance $\rr_1$ of $\wD$ in $\Deri$ is \emph{in the
   scope} of an instance $\rr_2$ of $\wD$ if $\rr_1$ introduced
   formulas that contain a free variable $x$ (i.e., $x$ occurs in a
   term in a predicate) and $\rr_2$ introduces the atom $x$ as a
   subformula (i.e. the propositional encoding of the binder $x$).  We
   can now permute the $\wD$-instances in $\DDeri$ such that whenever
   a rule instance $\rr_1$ depends on an instance $\rr_2$, then
   $\rr_2$ occurs below $\rr_1$ in $\DDeri$.
   We can now lift $\DDeri$ stepwise. First, observe that each line of
   in $\DDeri$ is $\PE\fequ$-equivalent to the propositional encoding $\PE P$ of a
   first-order formula $P$.  We now have to show
   that each other instance of $\wrD$ in $\DDeri$ is indeed the image of a correct
   application of $\wD$ or $\wfaD$ in first-order logic.
   If we have a $\wD$
   \begin{equation*}
     \vlinf{\wD}{}{\PE S\cons{x\vlor \PE A}}{\PE S\cons{\PE A}}
     \qquor
     \vlinf{\wD}{}{\PE S\cons{(x\vlor \PE B)\vlor \PE A}}{\PE S\cons{\PE A}}
   \end{equation*}
   then $x$ cannot occur freely in $A$, as otherwise the fibration
   property would be violated. We can therefore lift these instances to
   \begin{equation*}
     \vlinf{\wfaD}{}{S\cons{\forall x. A}}{S\cons{A}}
     \qquor
     \vlinf{\wD}{}{S\cons{(\forall x. B)\vlor A}}{S\cons{A}}
   \end{equation*}
   respectively.     
  %% A subformula of $\PE B$ is called \emph{weak} if it has been
  %% introduced by a weakening. All subformulas of a weak formula are
  %% also weak. Two weak subformulas $B'$ and $B''$ of $\PE B$ form a
  %% \emph{weak pair} if $\PE B\fequ\Scons{B'\vlor B''}$ for some context
  %% $\Sconhole$. We can assume without loss of generality that
  %% whenever weak subformulas $B'$ and $B''$ form a weak pair, they have
  %% been introduced by the same instance of $\wrD$ in
  %% $\PE\Deri_1$.\footnote{If $\PE\Deri_1$ is not of that shape, it can
  %% brought into this form by simple rule permutations, and then
  %% collapsing several weakenings into a single one.}  Now we show that
  %% $\PE\Deri_1$ can be lifted. For this, observe that whenever a
  %% weakening in $\PE\Deri_1$ deletes an atom $x\in\VAR$, it must also
  %% delete all atoms in the scope of the corresponding quantifier,
  %% because $\phi'$ is a fibration on the binding graph. Hence, each line
  If a  weakening happens inside a subformula $x\vlor \PE C$ or $x\vlan
  \PE C$ in $\DDeri$, then there are the following cases:
  \begin{equation*}%\small
    \scalebox{.9}{$
    \vlinf{\wD}{}{\PE S\cons{x\vlor\PE  D\vlor \PE C}}{\PE S\cons{x\vlor\PE  C}}
    \quad
    \vlinf{\wD}{}{\PE S\cons{x\vlan (\PE D\vlor \PE C)}}{\PE S\cons{x\vlan\PE  C}}
    \quad
    \vlinf{\wD}{}{\PE S\cons{(x\vlor\PE  D)\vlan \PE C}}{\PE S\cons{x\vlan \PE C}}
    $}
  \end{equation*}
  The first two cases can be lifted to
  \begin{equation*}
    \vlinf{\wD}{}{S\cons{\forall x.( D\vlor C)}}{S\cons{\forall x.  C}}
    \quand
     \vlinf{\wD}{}{S\cons{\exists x.( D\vlor C)}}{S\cons{\exists x.  C}}
  \end{equation*}
  respectively. But in the third case,  an $\exists$-quantifier would be
  transformed into an $\forall$-quantifier. But as $\phi$ has to
  preserve existentials, this third case cannot occur. All other situations can be lifted trivially, giving us
  $\vlder{\set{\wrD,\wfaD,\fequ}}{\Deri}{H}{G}$ as desired.
\end{proof}

@@

  \PE{(A\substof\phi)}$ and $\PE B_1$ are both propositional

@@

  propositional encodings, but the whole derivation cannot be

@@

  (and therefore a homomorphism) on the binding graphs, there must be

@@

  \item $\acDx$: We know that the premise of~\eqref{eq:acDeq} is a

@@

We first study the interactions between two rules. Certain cases are unsolved at
this stage, and they are considered later when we study the interactions between
two non-$\fequ$ rule instances separated by $\fequ$. Only non-trivial cases are
presented here:
\begin{itemize}
\item $\rr_1/\rr_2$, where $\rr_1 \in \set{\wrD, \wfaD}$ and $\rr_2 \in
\set{\acD, \cfaD, \me, \mfaD, \mexD}$:

@@

\item $\rr_1/\rr_2$, where $\rr_1 \in \set{\acD, \cfaD}$ and $\rr_2 \in
\set{\me, \mfaD, \mexD}$: 
  \begin{equation*}
  \scalebox{.9}{$
  \vlderivation{
    \vlin{\mfaD}{}{\Scons{\forall x.(A \vlor B)}}{
      \vlin{\cfaD}{}{\Scons{(\forall x.A) \vlor (\forall x.B)}}{
        \vlhy{\Scons{(\forall x.\forall x.A) \vlor (\forall x.B)}}}}}
  \leadsto
  \vlderivation{
    \vlin{\mfaD}{}{\Scons{\forall x.(A \vlor B)}}{
      \vlin{\fequ}{}{\Scons{(\forall x.A) \vlor (\forall x.B)}}{
        \vlin{\mfaD}{}{\Scons{\forall x.(\forall x.A \vlor B)}}{
          \vlhy{\Scons{(\forall x.\forall x.A) \vlor (\forall x.B)}}}}}}$}
  \end{equation*}

@@

\begin{equation*}
\scalebox{.9}{$
\vlderivation{
  \vlin{\fequ}{}{\forall y.\forall x.A}{
    \vlin{\wfaD}{(x \notin fv(\forall y.A))}{\forall x.\forall y.A}{
      \vlhy{\forall y.A}}}}
\leadsto
\vlderivation{
  \vlin{\wfaD}{(x \notin fv(A))}{\forall y.\forall x.A}{
    \vlhy{\forall y.A}}}$}
\end{equation*}

\begin{equation*}
\scalebox{.9}{$
\vlderivation{
  \vlin{\fequ}{}{\forall x.\forall y.A}{
    \vlin{\wfaD}{(x \notin fv(A))}{\forall y.\forall x.A}{
      \vlhy{\forall y.A}}}}
\leadsto
\vlderivation{
  \vlin{\wfaD}{(x \notin fv(\forall y.A))}{\forall x.\forall y.A}{
    \vlhy{\forall y.A}}}$}
\end{equation*}

@@

  \begin{equation*}
  \scalebox{.9}{$
  \vlderivation{
    \vlin{\me}{}{\forall x.((A \vlor B) \vlan (C \vlor D))}{
      \vlin{\fequ}{}{\forall x.((A \vlan C) \vlor (B \vlan
D))}{
        \vlin{\wrD}{}{(B \vlan D) \vlor (\forall x.(A \vlan C))}{
          \vlhy{B \vlan D}}}}}
  \leadsto
  \vlderivation{
    \vlin{\fequ}{}{\forall x.((A \vlor B) \vlan (C \vlor D))}{
      \vlin{\wrD}{}{\forall x.((B \vlor A) \vlan (D \vlor C))}{
        \vlin{\wrD}{}{\forall x.((B \vlor A) \vlan D)}{
          \vlin{\faD}{}{\forall x.(B \vlan D)}{
            \vlhy{B \vlan D}}}}}}$}
  \end{equation*}
  where in the second case, $x$ is free in $B \vlan D$.
